A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$

### 2025-04-05 paper_v02/article.tex
update latex table

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Universidade Federal da Bahia
   \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\Universidade Estadual\\de Feira de Santana}
\Plainauthor{Fernando Ferraz Ribeiro, Gilney Figueira Zebende}
%  and $DMC_x^2$
%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$ methods}
\Plaintitle{A Python/Zig optimized and customizable implementation for the Pdcca and DMCx2 methods}
\Shorttitle{\proglang{Python}/\proglang{Zig} implementation for the $\rho_{DCCA}$ and $DMC_x^2$}

@@

This paper presents the \pkg{Zebende} \proglang{Python} package, an implementation of the $DFA$, $DCCA$, $\rho_{DCCCA}$, $DMC_x^2$ and utility functions related to the methods. In section \ref{sec:calculations} the steps for calculating the $\rho_{DCCCA}$ and $DMC_x^2$ are presented and discussed. Section \ref{sec:optimization} shows how this library was implemented, the optimization technics and the recommended steps to use the library. In Section \ref{sec:results} the \pkg{Zebende} package is compared with other packages for \proglang{Python} and \proglang{R} that calculates the $\rho_{DCCA}$ in terms of performance and usability, leading to the conclusions in Section \ref{sec:summary}.

@@

None of the comparison packages can calculate multiple series comparison in the same input, so the $\rho_{DCCA}$ methods where wrapped around a loop that alternate the series used by the function, using a four two by two combination of the data indexes. The \pkg{DCCA-Python} package also didn't accepts multiple inputs for the time scale, so another loop was added in that test. The results are presented in the Table \ref{tab:time}.

\begin{table}[h]
  \label{tab:time}
  \centering
  \caption{Elapsed Time Comparison with four series}
  \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
    \hline
    Package & Implementation & Elapsed Time(s) & Percentage Increase \\
    \hline
    Zebende & Zig (opt) & 27.52  & - \\
    Zebende & Zig (basic) & 42.03 & 52.7\% \\
    DCCA & Python & 370.54 & 1239.4\% \\
    DCCA & R-Java & 1987.67 & 7202.4\% \\
    Zebende & Python & 2415.94 & 8803.4\% \\
    DFA & R & 20173.13 & 73,011.4\% \\
    \hline
  \end{tabular}
\end{table}

>>>>

%% - \author{} with primary affiliation (and optionally ORCID link)
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Fernando Ferraz Ribeiro~\orcidlink{0000-0002-0685-4774}\\Universidade Federal da Bahia
   \And Gilney Figueira Zebende~\orcidlink{0000-0003-2420-9805}\\Universidade Estadual\\de Feira de Santana}
\Plainauthor{Fernando Ferraz Ribeiro, Gilney Figueira Zebende}
%  and $DMC_x^2$
%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{A \proglang{Python}/\proglang{Zig} optimized and customizable implementation for the $\rho_{DCCA}$ and $DMC_x^2$}
\Plaintitle{A Python/Zig optimized and customizable implementation for the Pdcca and DMCx2 methods}
\Shorttitle{\proglang{Python}/\proglang{Zig} implementation for the $\rho_{DCCA}$ and $DMC_x^2$}

@@

This paper presents the \pkg{Zebende} \proglang{Python} package, an implementation of the $DFA$, $DCCA$, $\rho_{DCCCA}$, $DMC_x^2$ and utility functions related to the methods. In section \ref{sec:calculations} the steps for calculating the $\rho_{DCCA}$ and $DMC_x^2$ are presented and discussed. Section \ref{sec:optimization} shows how this library was implemented, the optimization technics and the recommended steps to use the library. In Section \ref{sec:results} the \pkg{Zebende} package is compared with other packages for \proglang{Python} and \proglang{R} that calculates the $\rho_{DCCA}$ in terms of performance and usability, leading to the conclusions in Section \ref{sec:summary}.

@@

None of the comparison packages can calculate multiple series comparison in the same input, so the $\rho_{DCCA}$ methods where wrapped around a loop that alternate the series used by the function, using a four two by two combination of the data indexes. The \pkg{DCCA-Python} package also didn't accepts multiple inputs for the time scale, so another loop was added in that test. The results are presented in the Table \ref{tab:time_4}.

  \begin{table}[h]

    \centering
    \caption{Elapsed Time Comparison with four series} \label{tab:time_4}
    \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
      \hline
      Package & Implementation & Elapsed Time(s) & Percentage Increase \\
      \hline
      Zebende & Zig (opt) & 35.88  & - \\
      Zebende & Zig (basic) & 59.74 & 52.7\% \\
      DCCA & Python & 447.67 &  1,243.49\% \\
      DCCA & R-Java & 2,206.3 & 6,126.17\% \\
      Zebende & Python & 3,076.23 & 8,542.83\% \\
      DFA & R & 28,818.016 & 80,049.42\% \\
      \hline
    \end{tabular}
  \end{table}

    \begin{table}[h]

      \centering
      \caption{Elapsed Time Comparison with twelve series} \label{tab:time_12}
      \begin{tabular}{@{}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{\hspace{1.0cm}}l@{}}
        \hline
        Package & Implementation & Elapsed Time(s) & Percentage Increase \\
        \hline
        Zebende & Zig (opt) & 120.55  & - \\
        Zebende & Zig (basic) & 213.08 & 76.64\% \\
        Zebende & Python & 3,761.40 & 3,085.47\% \\
        DCCA & Python & 4,675.17 &  3,871.65\% \\
        DCCA & R-Java & 4,8624.44 & 40,423.52\% \\
        DFA & R & 2,060,776.24 & 81,704,499.33\% \\
        \hline
      \end{tabular}
    \end{table}
